### 2025-06-05 ms/submission/AuthorComments-egusphere-2025-1046.qmd
address RC1.5: radiometric cps/concentration

---
title: "Response to Referees"
author: "Julien Vollering, on behalf of all co-authors"
format:
  html:
    embed-resources: true
toc: true
bibliography: ../ms.bib
---

```{=html}
<style>
.added {
  color: blue;
  font-weight: bold;
}
.deleted {
  color: red;
  text-decoration: line-through;
}
</style>
```

<https://www.soil-journal.net/peer_review/interactive_review_process.html>:

> Author comments (ACs) can only be posted by the contact authors of the manuscript in discussion on behalf of all co-authors. Co-authors can post CCs but not ACs. After the open discussion, during the final response, the author comments should be structured in a clear and easy-to-follow sequence: (1) comments from referees/community, (2) author's response, and (3) author's changes in manuscript.

# Response to Referee 1

<https://doi.org/10.5194/egusphere-2025-1046-RC1>

## General Comments

### Comment 1.1

> This article comprehensively addresses the modelling challenges of predicting peat depth from terrain variables. It takes high resolution terrain variables (derived from LiDAR) and low resolution (comparatively) airborne radiometric data across two individual peat landscapes in Norway and used Random Forest machine learning algorithm to train combinations of these variables to a multitude of peat depth probes and GPR peat depth measurements taken in order to establish the predict power of such variables for peat depth mapping.

### Comment 1.2

> Overall, I found the article to be generally well written, with a very comprehensive and detailed description of modelling mechanism, error derivation, and feature choice. It is a very long article, going into great detail in several areas, and below I suggest at least one section/topic that could be removed entirely to reduce length and increase overall readability. I suggest the authors review all sections for conciseness and reduce the article length where possible. The level of detail may mean a reader less familiar with machine learning modelling may find the article hard to follow. While I do find the article to be within the scope of SOIL, I would be concerned that it focuses heavily on the modelling methodology.

### Comment 1.3

> Additionally, there were several areas where the language used was quite casual for a scientific article. Several are highlighted under minor considerations below.

### Comment 1.4

> Finally, the main concern noted is the imbalance between the consideration given to radiometric data compared to terrain variables. Considering the title is stating terrain being “better” than radiometrics, and given the emerging understanding of the use of radiometric data in peat land mapping, there is are some fundamental errors in the methods presented, which may be biasing the conclusion alluded to in the title.

## Specific Comments

### Comment 1.5

> The first concern in the comparison of radiometric data to lidar terrain variables is related to the choice of radiometric variables. The authors opt to use potassium, uranium and thorium ground concentration units alongside the Total Count data from the full energy spectrum. These ground concentration measurements are derived from counts per second measurements on the airplane which are calibrated, usually using pads of known concentrations at a calibration facility. Therefore, concentration of any radioelement is a measurement of the concentration of that element in the top ~ 60 cm – 1 m of the soil. However, peat soils are different. Being organic, they don’t contain the typical geological material that make up soils. Therefore, they act as an attenuative environment to gamma rays. As the potential source of gamma rays in peat areas is blocked and attenuated by the peats, the concentration calibration is no longer physically valid. While these concentration data are indeed provided by the contractors of such surveys, it is now recognized that the counts per second measurement is a more appropriate unit when considering attenuation of gamma rays in peat soils (O'Leary et al, 2022, 2024). In particular considering depth, the deeper the peat, the greater the attenuation of gamma rays. Similarly, the wetter the peat the greater the attenuation of gamma rays. The use of concentration data is not valid for a study in predicting peat depth. I recommend either the authors convert these concentrations to counts per second, or remove all but the Total Count data from their analysis and consider take the next concern into account.

Thank you to the referee for pointing out this concern and providing a thorough explanation of their reasoning. We appreciate the references, which we have consulted.

After careful consideration, we disagree that the radiometrics variables we used are not valid for predicting peat depth. In particular, we assert that converting the concentrations we use to counts per second -- as the referee suggests -- would produce identical results. That is because the method we use to predict peat depth (Random Forest) is insensitive to any monotonic transformation of the predictors, like the scalar conversion between concentration and counts per second [see Table 10.1 on p.351 in @hastieElementsStatisticalLearning2009]. The Random Forest algorithm is based on decision trees, which partition the predictor space into regions based on the values of the predictors. As a result, any monotonic transformation of the predictors will not change the partitioning of the predictor space, and therefore will not change the predictions made by the model [see Fig. 9.2 on p.306 and Algorithm 15.1 on p.588 in @hastieElementsStatisticalLearning2009]. Appendix B3 in @baranwalHelicopterborneMagneticElectromagnetic2013 and Appendix A2 in @ofstadHelicopterborneMagneticRadiometric2015 confirm that the concentration values we use in our models are simply the counts per second multiplied by a constant conversion factor.

To try to avoid confusion on this point, we have added a short explanation in the Methods section that the distinction between counts per second and concentration is not relevant for our analysis and it does not affect the predictive power of the radiometric variables in our models.

::: {.callout appearance="simple"}
**In _Materials and methods_**

[Although radiometric data must be in units of counts per second to model attenuation directly [@olearyDigitalSoilMapping2022], we used the radiometric data as provided to us: in units of concentration for potassium, thorium, and uranium (converted from counts per second by scalar calibration factors).
The monotonic transformation between counts per second and concentration does not affect the tree-based machine learning algorithm that we used to model peat depth [@hastieElementsStatisticalLearning2009].]{.added} 
:::

### Comment 1.6

> There is an additional argument missing from within the authors discussion, namely the fact that we never know the initial source strength, or counts, of the gamma rays for a given footprint. The measurement at the airplane is an attenuated version of this initial source. This attenuation is controlled by the attenuation coefficient for a given element and depth, soil moisture, bulk density and porosity (Beamish 2013) of the peat soils. From a purely physics/modelling point of view, this makes the prediction of peat depth an underdetermined problem. Even if the soil moisture, bulk density and porosity was known absolutely, the initial source is never known and so any number of peat depths may result in any given gamma count at the airplane. Additionally, this modelling exercise cannot be performed on Total Count data as this is summed from the entire measurement energy spectrum, which contains multiple element specific attenuation coefficient, meaning the Total Count data is only ever indicative of attenuation variability across a site, with no ability to model anything quantitative. This puts Radiometrics in a natural disadvantage for a quantitative prediction of peat depth. Given the title of this article, I find that the discussion around the radiometrics lacked sufficient detail to make a fair comparison, which naturally results in such a bias towards terrain variables in such a modelling context. This is therefore not a result per say, but more a perfectly expected outcome. For this concern, I suggest a more comprehensive discussion around the physical limitations of radiometrics in the prediction of peat depth.

### Comment 1.7

> The main focus of this article is on the prediction of peat depth. However, the authors include several sections of the possibility of peatland extent mapping. This is not mentioned at all in the abstract, or the introduction in great detail. As this article is already quite long and complex, I suggest the removal of any sections and text related to mapping peat land extent as it is not the focus of the article and only acts to add unnecessary complexity to an already very technical methodology. The authors even mention in Line 535 that their aim was not to map extent. I suggest the removal of all reference to peatland extent prediction and instead focus on the prediction of peat depth. A much shorter reference to the importance of peatland extent knowledge could perhaps be mentioned in the conclusions, but a full analysis and discussion (section 4.1.4) is not appropriate in this article.

Thank you for this helpful suggestion. We initially thought it would be worth including the analysis on extent mapping---especially since our occurrence data at Skrimfjella is a good test set---but based on the comments of both referees we recognize that it distracts too much from the main focus. We have removed all sections related to mapping peatland extent (including section 4.1.4), to shorten the manuscript and improve its readability. We still discuss briefly the importance of coupling peat depth and extent mapping, since we consider this a research need directly related to the present study, but we have shortened this section as well.

## Technical Corrections

> Line 58- 59. There is no need to include this sentence with an example here, as the next paragraph goes into the necessary detail on Slope. This is an example of how the authors might reduce the overall size of the article.

> Section 2.2 I would suggest moving this opening paragraph to the end of this section as it acts more to sum up how the authors are using the various predictors. It mentions several of the predictors directly, but the are not described until later sections (for example 2.2.1). This would increase the readability of this section.

> Line 179: Remove “also using White Box” as this is obvious.

> Line 258: The authors mention density; however, they do not expand on this. Was this measured in the field, or an operator’s observation and subjective interpretation of density?

>  371: What is the relevance of the Persons correlation coefficient of 0.7. Was this tested at all? Readers may not be familiar with this so it should be explained a bit more.

> Table 2:  I recommend putting a vertical line between the results for both sites so as to easier distinguish between them.

> Heading Section 4.1.1 – “but not useless!” is very casual language to be using in a scientific article. This is just one example of this casual language. I suggest the authors review the article for this throughout.

> Line 464: “low hanging fruit” is also casual and colloquial and may not be understood by all cultures.

> Line 515: “large stocks really are large” – very vague and non-scientific comment. What is large?

> Line 530: remove the question at the start of this section.

> Line 549: This section is the first mention of the radiometric survey parameters. I would recommend moving some of this section to the Methods and Material section as it is useful descriptors of how the data came to be.

> Line 564: Typically, airborne radiometric surveys have strict conditions that they must fly under. One is related to rainfall occurrence and airborne surveys should not happen directly after rainfall. The authors statement with regards to air moisture should be clarified as otherwise the contractor may have been at fault by provide incorrect data, which would have implications for the usage of radiometric data in this area in this study.

> Line 628 – 629: remove the sentence starting “The rest of this section….” As it is unnecessary.

> Line 634: “luxury” is again a very casual phrasing within a scientific article.

> Line 664: “tricky” – casual

> Finally, there is no definite conclusion to this article, nor a heading stating same. I suggest the authors either add a section at the end and move some text here to highlight the main conclusions as currently the “discussion” section is quite large and probably not appropriate to act as a combined discussion and conclusion section.

# Response to Referee 2

<https://doi.org/10.5194/egusphere-2025-1046-RC2>

## General Comments

### Comment 2.1

> This article focuses on improving peat depth mapping in two distinct peatland landscapes in Norway using a digital soil mapping approach. The study employed the Random Forest (RF) algorithm to model both peat presence and depth, using high-resolution terrain data, remotely sensed radiometric data, and polygonised peat depth data from an existing map. The RF models were calibrated using field-measured data points collected across both regions, and variable importance was analysed.

### Comment 2.2

> Overall, this research addresses a significant challenge in peat depth mapping. The title is clear and likely to attract attention from readers interested in this topic. However, I have some concerns regarding the writing style. Certain sections use jargon excessively or are phrased in a way that feels less scientific, which may hinder clarity and accessibility for a broader audience.

## Specifics Comments

### Comment 2.3

> While the title is clear, I personally find it somewhat overconfident in summarising the results. This is mainly due to the relatively low performance of the RF models and the minimal difference observed between the terrain-only and terrain-plus-radiometric models. To better support the claims, it would be helpful to include a statistical test (e.g., a t-test) to assess whether the performance differences between models are statistically significant. Furthermore, incorporating a stand-alone model that uses only radiometric data as predictors would allow for a more balanced evaluation of the variable groups and help clarify the specific contribution of radiometric inputs in predicting peat depth in Norway.

Thank you for this suggestion. To clarify: the minimal marginal improvement in the terrain-plus-radiometric models compared to the terrain-only models is in fact evidence for the terrain variables being stronger predictors. Nevertheless, you are right that adding a radiometric-only model would provide a clearer comparison of the predictive contribution of radiometric versus terrain data. We now present models with all seven possible combinations of the three variable groups. These expanded results make clearer the predictive gap that we highlight in our title.

We have also added statistical tests to assess the significance of the differences in model performance. The results of these tests are now presented in tables in the appendix. The results show that the differences in model performance are statistically significant in most cases, which supports our claims about the relative importance of terrain variables in predicting peat depth. Specifically, 8 of the 12 direct comparisons between radiometric and terrain variables (Radiometric--Terrain and RadiometricDMK--TerrainDMK configuration pairs * 3 metrics * 2 sites) are statistically significant at the 0.05 level. We now describe these results in the Results section, to better support the claim in the title.

Some of the most important text revisions related to this comment are:

::: {.callout appearance="simple"}
**In _Methods_**

[We formally assessed the effect of predictor configuration on performance metrics using mixed-effects models to account for the cross-validation fold structure, and testing pairwise differences between configurations.]{.added}

**In _Results_**

[Compared to the *terrain* configuration, the *terrain + radiometric* configuration improved concordance correlation by 0.04, R^2^ by 0.04, and mean absolute error by 1 cm.]{.deleted}

[[For Skrimfjella] The *terrain* configuration outperformed the *radiometric* configuration by 0.27 in concordance correlation coefficient ($p < 0.001$; Appendix Table \@ref(tab:pairwiseSkrimCCC)), by 0.16 in R^2^ ($p = 0.190$; Appendix Table \@ref(tab:pairwiseSkrimRsq)), and by 10 cm in mean absolute error ($p = 0.038$; Appendix Table \@ref(tab:pairwiseSkrimMAE)).
]{.added}

[[For Ørskogfjellet] The *terrain* configuration outperformed the *radiometric* configuration by 0.24 in concordance correlation coefficient ($p = 0.068$; Appendix Table \@ref(tab:pairwiseOrskogCCC)), by 0.24 in R^2^ ($p = 0.004$; Appendix Table \@ref(tab:pairwiseOrskogRsq)), and by 25 cm in mean absolute error ($p = 0.086$; Appendix Table \@ref(tab:pairwiseOrskogMAE)).]{.added}

**In _Discussion_**

[Radiometric data had no predictive value at Ørskogfjellet, while at Skrimfjella they had minor influence in a relatively weak model.]{.deleted}

[Radiometric data had little to no predictive value at either site (poor performance of *radiometric* configuration), although they did contribute to the best model at Skrimfjella (marginal improvement in *all predictors* configuration compared to *terrain + DMK*).
These results contrast with earlier studies that found that radiometrics were useful predictors of peat depth [@keaneySpatialStatisticsEstimate2013; @gatisMappingUplandPeat2019; @kogantiMappingPeatDepth2023].]{.added}
:::

### Comment 2.4

> Figure 1: While the terrain-look map provides a general view of the study area, the relief of the regions remains unclear. Consider adding a clearer topographic representation (e.g., contour profile) to better illustrate landscape variation. Additionally, it would be helpful to explain the rationale for selecting Skrimfjella, which appears to have relatively limited peat coverage, especially given that the adjacent region seems to contain a larger peatland area.

Thank you for the suggestion to make the relief of the study areas clearer. We have added contour lines the maps of the study areas. We have not added contour profiles between specific points since we think that they may not be as useful for readers as the contour lines, but we are happy to add profiles if the contour lines are not enough.

We have also clarified the rationale for the delineation of the Skrimfjella study area, especially with respect to its limited peat coverage. In short: we were also investigating peatland extent mapping so low coverage was not disqualifying, and accessibility was also an important factor for this field work.

::: {.callout appearance="simple"}
**In _Materials and methods_**

At Skrimfjella we delineated a study area of 34 km^2^ based on radiometric coverage [(limiting to the west)]{.added} and accessibility [(limiting to the south), as part of a pilot project]{.added} (Fig. \@ref(fig:sites)b).
In Norway's AR5 national land cover dataset ["areal resources in scale 1:5000", @ahlstromAR5Klassifikasjonssystem2019], 1.5 km^2^ (4.5 %) of the study area is classified as 'mire' -- defined as areas with mire vegetation and at least 30 cm of peat depth.
[Relatively sparse peatland cover did not disqualify the area for our purposes, since we were also interested in peatland extent mapping in the pilot project.]{.added}
:::

### Comment 2.5

> It is unclear how the peat depth data are distributed, both statistically and spatially. While the sampling design is described in the text, I recommend including a figure showing the spatial distribution of the data points to help readers better understand the coverage and representativeness of the dataset. Additionally, a basic statistical summary (e.g., mean, median, range, standard deviation) of the peat depth values would be beneficial to provide context and a clearer picture of the conditions being modelled.

Thank you for this suggestion. We have added a figure showing the statistical and spatial distribution of the peat depths in the Results section. We have taken care to use the same map format as in the figure that presents land cover in the study areas, which we hope will help readers see the spatial representativeness of the dataset.

### Comment 2.6

> Since the radiometric data are originally at 50 m resolution and the terrain data are at 1 m resolution, the method used to resample these datasets to a common 10 m resolution could influence model performance. In particular, the use of cubic spline resampling for the radiometric data may affect how well its spatial variability is represented, especially when compared to the aggregated 10 m topographic data. I would be interested to hear your thoughts on how this resampling approach might have impacted the results.

### Comment 2.7

> I found the section on sampling design and peat depth measurement to be overly detailed for the main text. While this information is valuable, it may be more appropriate to move some of it to the appendix or supplementary material, as it is less central to the analysis and interpretation of results. This would help improve the flow and focus of the main manuscript.

### Comment 2.8

> In the model interpretation section, it appears that three different methods were used to assess variable importance: FIRM, permutation importance, and Shapley values. However, the implications of using these different approaches are not clearly discussed. Each method captures different aspects of variable influence and may lead to different interpretations. Could you clarify how the results from these methods align or differ, and what their respective implications are for understanding the drivers of peat depth in your study?

### Comment 2.9

> I recommend adding a dedicated section to summarise the key results and findings. This would help readers clearly see how the study addresses its original research aims and questions. A concise summary at the end of the results or discussion section would also improve the overall structure and coherence of the paper.

## Technical Corrections

> Table 2: I think you need to move this table somewhere after it is mentioned in the main text. At first, I was quite confused with this table. What does that percentage stand for?

We have moved the table after its first mention and adjusted the table caption to clarify that the percentages represent the proportion of the total number of 10 m cells across the stratifications. We have also added standard deviations to the mean depths, in accordance with comment 2.5.

> Line 394-397: I don’t quite get the point of this section. To my understanding, you compared between the quartile predictions and the observed data to see the prediction interval coverage probability. If so, I think it would be better to plot this prediction interval together with the observed data in scatter plot, like in Figure 3.

> Line 413-414: which curve? The information inside the parentheses is not clear to me.

> Figure 4: is the null derived from the calculation between observations and assumptions of 30 cm peat depth? Where is the standard error come from?

> Line 464: “low hanging fruit” ?

> Line 515: This sentence is unclear to me.

> Line 531: “0% in prediction” what does this sentence mean?

> Line 614: What are the ‘both indicators variables’ referring to?

# Response to Referee 3

### Comment 3.1

> Referee's comment goes here.

We thank the Referee for this insightful comment. We have revised the manuscript accordingly:

::: {.callout appearance="simple"}
**Original text:**

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.

**Revised text:**

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed do eiusmod tempor incididunt ut labore et **dolore magna aliqua** [and additional explanation of the concept]{.added}.
:::

### Comment 3.2

> Referee's comment goes here.

Thank you for highlighting this issue. We have made the following changes:

::: {.callout appearance="simple"}
**Original text:**

Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.

**Revised text:**

[We have completely revised this section]{.added} [to address the Referee's concerns.]{.added} [Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.]{.deleted}
:::

# Response to editorial support team

### Comment E.1

> Please add the section headlines "Appendix A" (in front of Table A1) and "Appendix B" (in front of Figure B1).