### 2025-08-29 ms/submission/AuthorComments2-egusphere-2025-1046.qmd
author response to acceptance

---
title: "Authors' response #2"
author: "Julien Vollering, on behalf of all co-authors"
format: pdf
---

We thank the topic editor for accepting our manuscript. We also thank the the reviewers again for their efforts.

We are not aware of any comments on our revised submission, or specific requested changes. The only change we have made in the text with respect to the revised submission is that we have included an in-text citation to our archived data, as requested in the email from the editorial support team. We have also added a new affiliation for one author.